Dynamics of Laterally Propagating Flames in X-ray Bursts. II. Realistic Burning \& Rotation

### 2021-01-20 xrb2/paper.tex
Updates to section 3.3

 \MarginPar{NF: I think we should plot peak enucdot as a function of time for each run, it would be useful for me to reference it in one of the later sections}

@@

Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with different initial temperature structures at $t = 70~\mathrm{ms}$. The hottest of the three plotted models, $\tt{F500\_3E8}$, has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}). {\color{blue} It also reaches slightly higher $\bar{A}$ values than the two cooler runs. We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front.}

\MarginPar{NF: the discussion here was originally comparing $\tt{F1000}$ and $\tt{F500\_3e8}$ using the incorrect figure caption, but it's actually comparing $\tt{F500_\_4e8}$ and $\tt{F1000}$. need to update?}
Similar to the $\bar{A}$ trends found for different rotation rates in Figure \ref{fig:abar}, the $\bar{A}$-$T$ phase plot in Figure~\ref{fig:abar_hot} suggests that the actual reactions happening in the hottest run may be different than for the cooler runs. Not only is the $\tt{F500\_3E8}$ flame itself significantly hotter overall, but there is a large causally connected region across a wide range of $\bar{A}$. This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures. The reaction pathways are also very disrupted for $\tt{F500\_3E8}$ compared to the cooler runs, suggesting that the hotter temperature leads to more turbulent mixing. Compared to the highest rotation rate run, $\tt{F1000}$, $\tt{F500\_3E8}$ reaches a peak temperature roughly $1 \times 10^9~\mathrm{K}$ higher ($\sim 2.7 \times 10^9~\mathrm{K}$ versus $\sim 1.7 \times 10^9~\mathrm{K}$). Though increasing the rotation rate does seem to concentrate the burning and increase the peak temperature, it does not reach the same level as directly increasing the temperature at the crust.

In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ is so hot that the organized flame structure is lost. The fuel layer undergoes stable burning and essentially simmers in place, rather than forming a flame. This process has been observed in neutron stars with accretion rates that exceed the Eddington limit \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model we have essentially recreated the conditions found in such stars with high accretion rates. This model burns so strongly that it is only run out to $40~\mathrm{ms}$. After an initial period of burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation. In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ across the domain and at heights up to $\sim 0.5\times 10^4~\mathrm{cm}$. There is still significant burning occurring even higher, with $\enucdot$ reaching $\sim 10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 10^4~\mathrm{cm}$. For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ and a maximum height of $\sim 0.5\times 10^4~\mathrm{cm}$, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed). Significant $\enucdot$ values for $\tt{F500\_3E8}$ are confined to the flame front, rather than spanning the entire domain.

\MarginPar{NF: should I include a plot for this or is it ok to just say it?}
Figure~\ref{fig:4e8_stacked_abar} shows $\bar{A}$ for $\tt{F500\_4E8}$. Again, burning extends across the domain and high into the atmosphere and fuel layer, lacking the characteristic flame structure shown in $\bar{A}$ plots for the lower temperature $500~\mathrm{Hz}$ runs (see Figure~\ref{fig:compare_500Hz_abar}). A distinct mass of material appears to have broken out of the atmosphere near the axis of symmetry. The atmosphere edge is located at $\sim 1.2\times 10^4~\mathrm{cm}$. {\color{blue}A similar effect is also visible (to a lesser extent) in the $\tt{F500\_3E8}$ plot in Figure \ref{fig:compare_500Hz_abar}, with a faint haze of material rising above the flame. However, this is likely to be a numerical artefact rather than a true physical effect, and a result of the boundary conditions imposed at the axis of symmetry for these simulations.} $\tt{F500\_4E8}$ clearly reaches much higher $\bar{A}$ values across the domain (especially at the latest snapshot, $t = 40~\mathrm{ms}$) compared to all the other runs described in this paper (see Figures \ref{fig:compare_500Hz_abar}, \ref{fig:time_series_500}, and \ref{fig:time_series_1000}). This is not surprising, since as we have shown in our simulations with differing initial temperature structures, the burning is very temperature-sensitive.

>>>>

@@

Figure~\ref{fig:compare_500Hz_abar} shows $\bar{A}$ for three $500~\mathrm{Hz}$ simulations with different initial temperature structures at $t = 70~\mathrm{ms}$. The hottest of the three plotted models, $\tt{F500\_3E8}$, has a faster propagating flame (this will be discussed further in Section \ref{ssec:temp_prop}). {\color{blue} It also reaches slightly higher $\bar{A}$ values than the two cooler runs. We do not plot $\tt{F500\_4E8}$ here because it fails to form a clear burning front. The Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plots of $\tt{F500}$ (left) and $\tt{F500\_3E8}$ (middle) also reflect these $\bar{A}$ features, with $\tt{F500\_3E8}$ reaching slightly higher $\bar{A}$ values. $\tt{F500\_3E8}$ also reaches higher $\enucdot$ values at the low end of the temperature range ($< 0.5 \times 10^9 \mathrm{K}$). There appear to be more causally connected regions across a range of $\bar{A}$ at low temperatures for $\tt{F500\_3E8}$ than for $\tt{F500}$, suggesting that the higher $\enucdot$ for $\tt{F500\_3E8}$ generates burning in certain reaction pathways that are not present in the cooler $\tt{F500}$ run.}
\MarginPar{relate this section back to Kiran's max temperature and $\enucdot$ plots once they're added, note that $\tt{F500\_3E8}$ doesn't reach as high temperatures, less evolved abar than $\tt{F1000}$}

In contrast to the models with $T_{\mathrm{star}} \leq 3 \times 10^8~\mathrm{K}$, $\tt{F500\_4E8}$ is so hot that the organized flame structure is lost. The fuel layer undergoes stable burning and essentially simmers in place, rather than forming a flame. This process has been observed in neutron stars with accretion rates that exceed the Eddington limit \citep{fujimoto1981,bildsten1998thermonuclear,keek2009effect}; in this model we have essentially recreated the conditions found in such stars with high accretion rates. This model burns so strongly that it is only run out to $40~\mathrm{ms}$. After an initial period when the burning moves across the domain, residual burning continues and eventually ignites the entire fuel layer at late times, as shown in Figures \ref{fig:4e8_stacked_enuc} and \ref{fig:4e8_stacked_abar} for three snapshots taken in the last ten seconds of the simulation. In Figure~\ref{fig:4e8_stacked_enuc}, $\enucdot$ reaches values of $10^{18} - 10^{20}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ across the domain and at heights up to $\sim 0.5\times 10^4~\mathrm{cm}$. There is still significant burning occurring even higher, with $\enucdot$ reaching $\sim 10^{15}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ at heights up to $\sim 0.8\times 10^4~\mathrm{cm}$. For comparison, the next hottest run ($\tt{F500\_3E8}$) only reaches maximum $\enucdot$ values on the order of $10^{18}~\mathrm{erg}\,\mathrm{g}^{-1} \mathrm{s}^{-1}$ {\color{blue} (see Figure ?? Kiran $\enucdot$ plot)}, even at the latest timesteps ($\sim 70~\mathrm{ms}$, when the flame is most developed). {\color{blue} Significant $\enucdot$ values for all other runs are confined to the flame front, rather than spanning the entire domain}.

Figure~\ref{fig:4e8_stacked_abar} shows $\bar{A}$ for $\tt{F500\_4E8}$. Again, burning extends across the domain and high into the atmosphere and fuel layer, lacking the characteristic flame structure shown in $\bar{A}$ plots for the lower temperature $500~\mathrm{Hz}$ runs (see Figure~\ref{fig:compare_500Hz_abar}). A distinct mass of material appears to have broken out of the atmosphere near the axis of symmetry. The atmosphere edge is located at $\sim 1.2\times 10^4~\mathrm{cm}$. {\color{blue}A similar effect is also visible (to a lesser extent) in the $\tt{F500\_3E8}$ plot in Figure \ref{fig:compare_500Hz_abar}, with a faint haze of material rising above the flame. However, this is likely to be a numerical artefact rather than a true physical effect, and a result of the boundary conditions imposed at the axis of symmetry for these simulations.} $\tt{F500\_4E8}$ clearly reaches much higher $\bar{A}$ values across the domain (especially at the latest snapshot, $t = 40~\mathrm{ms}$) compared to all the other runs described in this paper (see Figures \ref{fig:compare_500Hz_abar}, \ref{fig:time_series_500}, and \ref{fig:time_series_1000}). This is not surprising, since as we have shown in our simulations with differing initial temperature structures, the burning is very temperature-sensitive. \MarginPar{is it generally temperature-sensitive, or only sensitive to the highest temperatures?}

 {\color{blue} Though $\tt{F500\_4E8}$ does not form a distinct burning front, it does achieve greatly enhanced burning, as shown in the Figure~\ref{fig:abar_hot} $\bar{A}$-$T$ phase plot (right panel). Not only is $\tt{F500\_4E8}$ significantly hotter overall compared to all other runs}, but there is a large causally connected region across a wide range of $\bar{A}$. This indicates that the hotter temperature has facilitated significant burning in reaction pathways that were not favored at the lower temperatures. The reaction pathways are also very disrupted for $\tt{F500\_4E8}$ compared to the cooler runs, suggesting that the hotter temperature leads to more vigorous mixing. {\color{blue}Indeed, this appears to be the case looking at Figure~\ref{fig:4e8_stacked_abar} and the right panel of Figure~\ref{fig:uv_hot} (discussed further in section~\ref{ssec:temp_prop}). The disrupted reaction pathways resemble those found in Figure \ref{fig:abar} for the highest rotation $\tt{F1000}$ run, though they are even more dramatically disrupted for $\tt{F500\_4E8}$. $\tt{F500\_4E8}$ is also noticably hotter than $\tt{F1000}$, even though it is run for significantly less time ($t = 40~\mathrm{ms}$ vs $t = 100~\mathrm{ms}$). While the correlation between increased temperature, greater mixing, and enhanced burning seen in $\tt{F500\_4E8}$ agrees with our previous anaylses, without an increased rotation and Coriolis force the $\tt{F500\_4E8}$ run seems insufficiently confined to maintain a burning front.} \MarginPar{is this interpretation of the lack of Coriolis force correct?}
